\chapter{Acknowledgements}

First and foremost, thanks to Peter Keightley. He has been an excellent supervisor and mentor for the last four years. Peter's help and guidance has helped me develop in many ways and has made my PhD an incredibly enjoyable and rewarding experience. Peter's thorough and rigorous approach to science is something that I aspire to. Thanks for introducing me to disc golf too!

I would also thank Brian Charlesworth for helping me understand the thornier aspects of population genetics that I have come across. Brian has been very patient with me and willing to give help and advice whenever I have asked for it, which has been hugely helpful throughout my PhD.

Thanks to Deborah Charlesworth for being very kind and generous in helping me understand many aspects of evolutionary biology, not limited to my own research. Thanks to participants of the evolutionary genetics lab group and genetics journal club in IEB for great discussions and feedback about my research, particularly Susie Johnston and Konrad Lohse. I would also extend massive thanks to Sally Otto and members of her lab group at UBC for giving me such a welcoming place to work and write-up my thesis in Vancouver.

Thanks to members of the Keightley lab past and present for listening to me go on and on about selective sweeps or other things for the past four years. In no particular order: thanks Ben Jackson, Rory Craig, Susanne Kraemer, Eva Deinum, Thanasis Kousathanas and Matty Hartfield. Rob Ness taught me to code and helped me out in numerous ways, it's just a shame he left Edinburgh when he did. Without the help of Dan Halligan, who tolerated me bugging him even after he left the lab, I would have been very lost.

On the personal side of things, my family has been very supportive throughout my whole PhD, particularly my amazing Mum and Dad. My brothers and their SAPs (spouses and partners) are all really great, but Stella, you really nailed it. Thanks to my friends in IEB and the whole Berwickshire gang for palling around with me in Edinburgh. Thanks to Jaz and Michael for all the great hangs in Vancouver. 

Finally, Arya has looked after me and kept my head straight for the past four years. She is the best and even married me during this PhD, can you believe that?
