\chapter{Declaration}

\noindent
I declare that the thesis has been composed by myself and that the work has not be
submitted for any other degree or professional qualification. I confirm that the work
submitted is my own, except where work which has formed part of jointly-authored
publications has been included. My contribution and those of the other authors to this
work have been explicitly indicated below. I confirm that appropriate credit has been
given within this thesis where reference has been made to the work of others.

\noindent
I composed this thesis, under guidance from Professor Peter Keightley, and with contributions from other authors for individual chapters as stated below. Parts of the introduction were included in a published review paper. Chapter 2 is published and Chapter 3 is a manuscript currently under review. I have retained the use of "we" in Chapters 2-3 and, for consistency, also use ``we" in Chapter 4. I use ``I" for Chapter 5. All published papers are reproduced in the Appendix. I confirm that I carried out all analyses myself.

\noindent
\textbf{Chapter 2}
\noindent
I designed the analyses with Professor Peter Keightley and Dr Rob Ness. I performed all analyses and wrote the paper, under guidance from Professor Keightley and Dr Ness.

\noindent
\textbf{Chapter 3}
\noindent
I designed the analyses for this project with Professor Peter Keightley. I performed all analyses and wrote the paper, under guidance from Professor Keightley.

\noindent
\textbf{Chapter 4}
\noindent
I designed the analyses for this project with Professor Peter Keightley and Professor Brian Charlesworth. I performed all analyses and wrote the paper, under guidance from Professors Keightley and Charlesworth.
